\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-12-17 archive.tex
progress on DSL figure

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\denotation}[1]{{\llbracket #1 \rrbracket}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
\newcommand{\scode}[1]{{\tiny\texttt{#1}}}
\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
% Use the following line for the initial blind version submitted for review:
%\usepackage[nohyperref]{icml2018}

@@

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet,calc}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzset{squiggle/.style={decorate, decoration={snake,amplitude=.4mm}}}
\usepackage{dsfont}
\usepackage{amsmath}

@@

\begin{figure}
  \begin{tikzpicture}[line width=0.4mm]
    \draw (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
    \node at (5.5,1.25) {\textsc{\textbf{Wake}}};

@@

    \node at (0.5,-4.75) {\textbf{\textsc{Memory Consolidation}}};
    \draw (-2,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-2,-13) -- (-2,-5);
    \node at (11.5,-4.75) {\textbf{\textsc{Dreaming}}};
    \draw (14,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (14,-13) -- (14,-5);

    \begin{scope}[shift={(9,-4.5)}]
      \node(dreaming) at (2,-1) {\underline{Fantasies}};
      \node at ([yshift=-0.75cm,xshift=-1.5cm]dreaming.south) (d){DSL};
      \node at ([xshift=1.5cm]d.east) (p1){prog.};
      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
        \NeuralNetwork{0.17}};
      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
      \draw [dashed,red] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);

      \node(replay) at (2,-5) {\underline{Replays}};
      \node[align=center] at ([yshift=-0.75cm,xshift=-1.5cm]replay.south) (d){progs.\\for\\task};
      \node at ([xshift=1.5cm]d.east) (p1){prog.};
      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
        \NeuralNetwork{0.17}};
      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
      \draw [dashed,red] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);

@@

    %%}b memory consolidation
    \begin{scope}[shift={(-2,-4.5)}]
      \node[align=center] at (0.8,-1.2) (f1){progs. for\\task$_1$:\\\scode{(+ (car z) 1)}};
      \node[align=center] at ([xshift = 1cm]f1.east) (f2){progs. for\\task$_2$:\\\scode{(cons (+ 1 1))}};
      \node[align=center] at ([yshift = 0cm,xshift=1cm]f2.east) (f3){\textbf{DSL}\\\scode{cons}, \scode{+}, \\...};
      \node(c)[rectangle, rounded corners, dashed, draw, minimum width = 3.75cm, minimum height = 3cm, anchor = north] at ([yshift=-0.8cm]f2.south) {};
      \draw [-> ] (f1.south) -- (f1.south|-c.north);
      \draw [-> ] (f2.south) -- (f2.south|-c.north);
      \draw [-> ] (f3.south) -- (f3.south|-c.north);

@@

@@

This dream-learning architecture unifies two lines of prior work.
One line of work
considers the problem of learning new concepts, abstractions, or
`options' from
experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
while the other line of work considers the problem of learning how to
deploy those concepts
efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
goal with DreamCoder is to show that the combination of these ideas is
uniquely powerful, and pushes us toward program-writing systems that,
like human learners, autonomously acquire the expertise needed to
navigate a new domain of problems.

@@

\system~takes as its goal to
acquire domain-specific expertise,
which it learns by
solving a collection of
programming \textbf{tasks} -- typically about a hundred.
During its Wake phase,
it searches for programs solving tasks,
guided by its current DSL and recognition model.
It cycles through Wake

@@

\textbf{Idea 2:} Rather than explicitly enumerate every refactoring,
we developed a technique for compactly representing and efficiently
calculating the set of all refactorings.  The key idea is to represent
the set of refactorings using a \emph{version
  space}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}.
A version space is a tree-shaped data structure
that compactly represents a large set of programs.
We first define version spaces over $\lambda$-calculus expressions,
and then give a dynamic program for constructing a version space
that represents the set of possible refactorings (i.e., $R_n(p)$).
This technique is astronomically more efficient than
explicitly representing
the space of possible refactorings:
for the example in Figure~\ref{mapFactor},
we represent
the space of refactorings using a version space with
$10^6$ nodes,
which encodes $10^{14}$ refactorings.
Formally, a version space is either:
\begin{itemize}
\item A deBuijn\footnote{deBuijn indices are an alternative way of naming variables in $\lambda$-calculus. When using deBuijn indices, $\lambda$-abstractions are written \emph{without} a variable name, and variables are written as the count of the number of $\lambda$-abstractions up in the syntax tree the variable is bound to. For example, $\lambda x.\lambda y. (x\;y)$ is written $\lambda\lambda (\$1\;\$0)$ using deBuijn indices. See~\cite{pierce} for more details.} index: written $\$i$, where $i$ is a natural number
\item An abstraction: written $\lambda v$, where $v$ is a version space
\item An  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
\item A  union: $\uplus V$, where $V$ is a set of version spaces
\item The empty set, $\varnothing$
\item The set of all $\lambda$-calculus expressions, $\Lambda$
\end{itemize}
The purpose of a version space to compactly represent a set of programs.
We refer to this set as the \textbf{extension} of the version space:
\begin{definition}
  The \textbf{extension} of a version space $v$ is written $\denotation{v}$
  and is defined recursively as:
  \begin{align*}
    \denotation{\$i}& = \left\{\$i \right\}\\
    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}\\
    \denotation{(v_1\; v_2)}& = \left\{(e_1\;e_2) : e_1\in \denotation{v_1},\;e_2\in \denotation{v_2} \right\}\\
    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}\\
    \denotation{\varnothing}& = \varnothing\\
    \denotation{\Lambda}& = \Lambda
    \end{align*}
\end{definition}
Version spaces also support efficient membership checking, which we
write as $e\in v$, and which is equivalent to $e\in \denotation{v}$.
Important for our purposes, it is also efficient to refactor the
members of a version space's extension in terms of a new DSL.
We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
\begin{align}
  \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
    e\text{, if $e\in \mathcal{D}$ and $e\in v$. Exploits the fact that $e\in v$ can be efficiently computed.}\\
    \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
  \end{cases}\label{FACTORING}\\\nonumber
  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}\\\nonumber
  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\\nonumber
  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})\\\nonumber
  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\text{size}(e|\mathcal{D})\nonumber
  \end{align*}
where $\text{size}(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
is the size of the syntax tree of $e$,
when members of $\mathcal{D}$ are counted as having size 1.
Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\text{size}(p|\mathcal{D})$.

>>>>

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\systemEnding}{\textsc{DreamCoder}}
\newcommand{\lowerBound}{\mathscr{L}}
\newcommand{\denotation}[1]{{\llbracket #1 \rrbracket}}
\newcommand{\code}[1]{{\footnotesize\texttt{#1}}}
\newcommand{\scode}[1]{{\tiny\texttt{#1}}}
\newcommand{\mcode}[1]{{\scriptsize\texttt{#1}}}
\newcommand{\codechar}[1]{{\footnotesize{\texttt{"#1"}}}}
% Use the following line for the initial blind version submitted for review:
%\usepackage[nohyperref]{icml2018}

@@

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet,calc,tikzmark}
\usetikzlibrary{arrows.meta}
\usetikzlibrary{positioning}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzset{squiggle/.style={decorate, decoration={snake,amplitude=.4mm}}}
\usepackage{dsfont}
\usepackage{amsmath}

@@

\begin{figure}
  \begin{tikzpicture}[line width=0.4mm]
    \draw[fill=white] (-1,1) -- (13,1) -- (13,-4) -- (-1,-4) -- (-1,1);
    \node at (5.5,1.25) {\textsc{\textbf{Wake}}};

@@

    \node at (0.5,-4.75) {\textbf{\textsc{Sleep: Consolidation}}};
    \draw (-2,-5) -- (3,-5) -- (3,-10) -- (5.5,-10) -- (5.5,-13) -- (-2,-13) -- (-2,-5);
    \node at (11.5,-4.75) {\textbf{\textsc{Sleep: Dreaming}}};
    \draw (14,-5) -- (9,-5) -- (9,-10) -- (6.5,-10) -- (6.5,-13) -- (14,-13) -- (14,-5);

    \begin{scope}[shift={(9,-4.5)}]
      \node(dreaming) at (2,-1) {\underline{Fantasies}};
      \node at ([yshift=-0.75cm,xshift=-1.5cm]dreaming.south) (d){DSL};
      \node at ([xshift=1.5cm]d.east) (p1){prog.};
      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
        \NeuralNetwork{0.17}};
      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
      \draw [dashed] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);

      \node(replay) at (2,-5) {\underline{Replays}};
      \node[align=center] at ([yshift=-0.75cm,xshift=-1.5cm]replay.south) (d){progs.\\for\\task};
      \node at ([xshift=1.5cm]d.east) (p1){prog.};
      \draw[squiggle,-> ] (d.east) -- node[sloped,above]{\small sample} (p1.west);
      \node at ([xshift = 1.5cm]p1.east) (t1){ task};
      \draw [-> ] (p1.east) -- node[above]{\small run} (t1.west);
      \node(n) at ([yshift=-1.2cm,xshift=0.4cm]p1.south) {
        \NeuralNetwork{0.17}};
      \draw [->,red] (t1.south) to[out = -90,in = 0]  ([xshift=0.4cm]n.east);
      \draw [dashed] (p1.south) to[out=-120,in=180] node[above,fill=white]{\color{black}Loss} ([xshift=-0.4cm]n.west);

@@

    %%}b memory consolidation
    \begin{scope}[shift={(-2,-4.5)}]
      \node[align=center] at (0.8,-1.2) (f1){progs. for\\task$_1$:\\\mcode{(+ (car z) 1)}};
      \node[align=center] at ([xshift = 1cm]f1.east) (f2){progs. for\\task$_2$:\\\mcode{(cons (+ 1 1))}};
      \node[align=center] at ([yshift = 0cm,xshift=0.4cm]f2.east) (f3){\textbf{DSL}\\\mcode{cons}, \mcode{+}, \\...};
      \node(c)[rectangle, rounded corners, dashed, draw, minimum width = 3.75cm, minimum height = 3cm, anchor = north] at ([yshift=-0.8cm,xshift=-0.3cm]f2.south) {};
      \draw [-> ] (f1.south) -- (f1.south|-c.north);
      \draw [-> ] (f2.south) -- (f2.south|-c.north);
      \draw [-> ] (f3.south) -- (f3.south|-c.north);

@@

\begin{figure}
  \begin{tikzpicture}
%    \draw (0,0) rectangle (10,5);
    \node(t1)[align=left] at (1,-1) {
      \small +106 769-43$\to $106.769.43\\
      \small +83 973-831$\to $83.973.831\\
      \small $f(\text{\code{s}}) = $\code{(}\subnode{f01}{$f_0$}\code{  \codechar{.} \codechar{-}}\\
      \small \hspace{1.25cm}\code{($f_0$ \codechar{.} \codechar{ }}\\
      \small \hspace{1.4cm}\code{(cdr s)))}
    };
    \node(t2)[align=left] at ([xshift=2cm]t1.east) {
      \small Dream Coder$\to$Dr. Dream\\
      \small Andy Cencici$\to$Dr. Andy\\
      \small Jan Kotas$\to$Dr. Jan\\
      \small Nancy Jones$\to$Dr. Nancy\\
      $f(\text{\code{s}}) = $\code{($f_3$ ($f_4$ \codechar{ } s))}
    };
    \node(t3)[align=left] at ([xshift=2cm]t2.east) {
      \small Dream (Coder)$\to$Coder\\
      \small 1 (2) 3 4$\to$2\\
      \small the (dog) ran$\to$ran\\
      \small +817(33)99$\to$33\\
      $f(\text{\code{s}}) = $\code{($f_6$ \codechar{(} \codechar{)} s)}
    };

    \node(substitution)[draw,rounded corners,align=center] at (1,-4) {
      \small $f_0($\code{s}$,$\code{a}$,$\code{b}$) \,=\, $\\
       \mcode{(map ($\lambda$ (x) (if}\\
      \mcode{(= x a) b x)) s)}
    };
    \node[anchor = north] at (substitution.south) {substitute characters};
    \node(cat)[draw,rounded corners,align=center] at (1,-6) {
      \small $f_1($\code{a}$,$\code{b}$) \,=\, $\\
       \mcode{(fold ($\lambda$ (x y)}\\\mcode{ (cons x y)) a b))}
    };
    \node[anchor = north] at (cat.south) {concat strings};
    \node(looping)[draw,rounded corners,align=center] at (1,-8) {
      \small $f_2($\code{c}$,$\code{p}$,$\code{s}$) \,=\, $\\
       \mcode{(unfold l ($\lambda$ (x)}\\\mcode{(= c (car x)))}\\\mcode{ p ($\lambda$ (z) (cdr z)))}
    };
    \node[anchor = north] at (looping.south) {loop until char};

    \node(constant)[draw,rounded corners,align=center] at (4.5,-4) {
      \small $f_3($\code{s}$) \,=\, $\\
       \mcode{($f_0$ s string)}
    };
    \node[anchor = north] at (constant.south) {prepend constant};
    \node(drop)[draw,rounded corners,align=center] at (4.5,-6) {
      \small $f_4($\code{c}$) \,=\, $\\
      \mcode{($f_2$ c ($\lambda$ }\\\mcode{(z) (car z)))}
    };
    \node[anchor = north] at (drop.south) {prefix};
    \node(suffix)[draw,rounded corners,align=center] at ([yshift=-2cm]drop.center) {
      \small $f_5($\code{s}$,$\code{c}$) \,=\, $\\
      \mcode{(fold ($\lambda$ (x y) }\\\mcode{(cdr (if (= x c)}\\\mcode{ s y))) s s)}
    };
    \node[anchor = north] at (suffix.south) {suffix};
    \node(delimited)[draw, rounded corners, align=center,text width=2.5cm] at ([xshift=3.5cm]constant.center) {
      \code{$f_6($x$,$y$,$s$)\,=\,$}\\
      \mcode{($f_4$ x ($f_5$ y c))}};
    \node[anchor = north] at (delimited.south) {delimited substring};
    \draw[->] (delimited.west) -- (drop.east);
    \draw[->] (delimited.west) -- (suffix.east);

    \draw[->] (drop.west) -- (looping.east);
    \draw[->] (constant.west) -- (cat.east);

    \draw[->] ([xshift=-0.5cm]t2.south) -- (constant.north);
    \draw[->] ([xshift=0.3cm]t2.south) to[in=30,out=-30] ([xshift=1cm]drop.north);

    \draw[->] ([yshift=0.5cm,xshift=-0.4cm]t1.south) to[in=100,out=200] (substitution.north);
    \draw[->] ([yshift=0.8cm,xshift=-0.5cm]t1.south) to[in=120,out=200] ([xshift=-0.5cm]substitution.north);
    \draw[->] ([xshift=-0.5cm]t3.south) -- (delimited.north);

@@

This dream-learning architecture brings together two lines of prior work.
One line of work
considers the problem of learning new concepts, abstractions, or
`options' from
experience~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,solomonoff1989system,DBLP:journals/corr/abs-1110-5667,stolle2002learning},
while the other line of work considers the problem of learning how to
deploy those concepts
efficiently~\cite{devlin2017robustfill,balog2016deepcoder,NGDS}.  Our
goal with DreamCoder is to show that the combination of these ideas is
uniquely powerful, and pushes us toward program-writing systems that,
like human learners, autonomously acquire the expertise needed to
navigate a new domain of problems.

@@

\system~takes as its goal to acquire domain-specific expertise, which
it learns by solving a collection of programming \textbf{tasks} --
typically about a hundred.  It alternatingly finds programs that solve
tasks (Wake -- Figure~\ref{threeCycles} top); improving its DSL by
discovering and reusing domain-specific subroutines (Memory
Consolidation -- Figure~\ref{threeCycles} left); and training a neural
network that efficiently guides search for programs in the DSL
(Dreaming -- Figure~\ref{threeCycles} right).  The learned DSL
effectively encodes a prior on programs likely to solve tasks in the
domain, while the neural net looks at the example input-output pairs
for a specific task and produces a ``posterior'' for programs likely
to solve that specific task (Figure~\ref{threeCycles} middle).  The
neural network thus functions as a \textbf{recognition model}
supporting a form of approximate Bayesian program induction, jointly
trained with a \textbf{generative model} for programs encoded in the
DSL, in the spirit of the Helmholtz machine~\cite{hinton1995wake}. The
recognition model ensures that searching for programs remains
tractable even as the DSL (and hence the search space for programs)
expands.

@@

\textbf{Idea 2:} Rather than explicitly enumerate every refactoring,
we developed a technique for compactly representing and efficiently
calculating the set of all refactorings.  The key idea is to represent
the set of refactorings using a \emph{version
  space}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}.
A version space is a tree-shaped data structure
that compactly represents a large set of programs.
We first define version spaces over $\lambda$-calculus expressions,
and then give a dynamic program for constructing a version space
that represents the set of possible refactorings (i.e., $R_n(p)$).
This technique is astronomically more efficient than
explicitly representing
the space of possible refactorings:
for the example in Figure~\ref{mapFactor},
we represent
the space of refactorings using a version space with
$10^6$ nodes,
which encodes $10^{14}$ refactorings.
Formally, a version space is either:
\begin{itemize}
\item A deBuijn\footnote{deBuijn indices are an alternative way of naming variables in $\lambda$-calculus. When using deBuijn indices, $\lambda$-abstractions are written \emph{without} a variable name, and variables are written as the count of the number of $\lambda$-abstractions up in the syntax tree the variable is bound to. For example, $\lambda x.\lambda y. (x\;y)$ is written $\lambda\lambda (\$1\;\$0)$ using deBuijn indices. See~\cite{pierce} for more details.} index: written $\$i$, where $i$ is a natural number
\item An abstraction: written $\lambda v$, where $v$ is a version space
\item An  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
\item A  union: $\uplus V$, where $V$ is a set of version spaces
\item The empty set, $\varnothing$
\item The set of all $\lambda$-calculus expressions, $\Lambda$
\end{itemize}
The purpose of a version space to compactly represent a set of programs.
We refer to this set as the \textbf{extension} of the version space:
\begin{definition}
  The \textbf{extension} of a version space $v$ is written $\denotation{v}$
  and is defined recursively as:
  \begin{align*}
    \denotation{\$i}& = \left\{\$i \right\}&
    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}&
    \denotation{(v_1\; v_2)}& = \left\{(e_1\;e_2) : e_1\in \denotation{v_1},\;e_2\in \denotation{v_2} \right\}\\
    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}&
    \denotation{\varnothing}& = \varnothing&
    \denotation{\Lambda}& = \Lambda
    \end{align*}
\end{definition}
Version spaces also support efficient membership checking, which we
write as  $e\in \denotation{v}$.
Important for our purposes, it is also efficient to refactor the
members of a version space's extension in terms of a new DSL.
We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
\begin{align}
  \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
    e\text{, if $e\in \mathcal{D}$ and $e\in \denotation{v}$. Exploits the fact that $\denotation{e}\in v$ can be efficiently computed.}\\
    \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
  \end{cases}\label{FACTORING}\\\nonumber    
  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}&%\\\nonumber
  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\\nonumber
  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})&%\\\nonumber
  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\text{size}(e|\mathcal{D})\nonumber
  \end{align*}
where $\text{size}(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
is the size of the syntax tree of $e$,
when members of $\mathcal{D}$ are counted as having size 1.
Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\text{size}(p|\mathcal{D})$.
